---
title: "Comparison of temperature from different sources"
authors:
  - bangyou-zheng
date: '2023-03-09'
slug: compare-temperature
tags:
  - R
  - climate
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE,
                      out.height = 600, out.width = "100%")
suppressPackageStartupMessages(library(tidyverse))
```

```{r read-data}
records <- readRDS("../compare-radiation/weather-records.Rds")
records <- records %>% 
    mutate(Source = factor(Source, 
                           levels = c("agcd", "power", "ppd", "drill")))
```


## Data source

The weather records are sourced for 79 sites from 

* [SILO Patched Point Dataset](https://www.longpaddock.qld.gov.au/silo/point-data/) with label `ppd`
* [SILO Data Drill](https://www.longpaddock.qld.gov.au/silo/gridded-data/) with label `drill`
* [Australian Gridded Climate Data](http://www.bom.gov.au/climate/austmaps/about-agcd-maps.shtml) from BOM with label `agcd`
* [NASA POWER](https://power.larc.nasa.gov/) with label `power`

The daily radiations (MJ/m^2) are downloaded from 1990-01-01 to 2022-12-31.

## Sites

There are 79 sites sourced from Lilley et al (2019) across Australian grain belt.

```{r fig-site, results='asis', out.height = 400, out.width = "100%"}
sites <- read.csv("../compare-radiation/sites.csv", as.is = TRUE)


library(leaflet)

m <- leaflet() %>%
    addTiles() %>%  # Add default OpenStreetMap map tiles
    addMarkers(lng=~Longitude, lat=~Latitude, popup=~Site, data = sites) %>%
    leaflet.extras::suspendScroll(hoverToWake = FALSE)
m

```



## Yearly average minimum temperature

I calculated the yearly average minimum temperature and then anomaly of average minimum temperature from 1990 to 2022.

```{r mint-abnormal-year, fig.height=6, fig.cap="Anomaly of average minimum temperature (C)"}


yearly_records <- records %>%  
    filter(mint > -100) %>% 
    group_by(Source, Number, year) %>% 
    summarise(mint = mean(mint), .groups = "drop") %>% 
    group_by(Source, Number) %>% 
    mutate(abnormal = mint - mean(mint)) %>%
    ungroup() 
yearly_records %>% 
    ggplot(aes(year, abnormal, color = factor(Number))) +
    geom_line() +
    guides(color = "none") +
    facet_wrap(~Source) +
    geom_hline(yintercept = 0) +
    theme_bw() +
    ylab("Anomaly of average minimum temperature (C)") +
    xlab("Year")

```


```{r fig-yearly-mint-compare, out.height=200, fig.cap="Comparison of yearly average minimum temperature between AGDC and other sources"}


yearly_records %>% 
    filter(Source == "agcd") %>% 
    select(Number, year, base = mint) %>% 
    right_join(yearly_records %>% 
                   filter(Source != "agcd") %>% 
                   select(Number, Source, year, mint)) %>% 
    ggplot() +
    geom_point(aes(base, mint)) +
    facet_wrap(~Source) +
    theme_bw() +
    geom_abline() +
    xlab("Yearly average minimum temperature from AGCD (C)") +
    ylab("Yearly average minimum temperature from other sources (C)")

```



## Yearly average maximum temperature

I calculated the yearly average maximum temperature and then anomaly of average maximum temperature from 1990 to 2022.

```{r maxt-abnormal-year, fig.height=6, fig.cap="Anomaly of average maximum temperature (C)"}


yearly_records <- records %>%  
    filter(maxt > -100) %>% 
    group_by(Source, Number, year) %>% 
    summarise(maxt = mean(maxt), .groups = "drop") %>% 
    group_by(Source, Number) %>% 
    mutate(abnormal = maxt - mean(maxt)) %>%
    ungroup() 
yearly_records %>% 
    ggplot(aes(year, abnormal, color = factor(Number))) +
    geom_line() +
    guides(color = "none") +
    facet_wrap(~Source) +
    geom_hline(yintercept = 0) +
    theme_bw() +
    ylab("Anomaly of average maximum temperature (C)") +
    xlab("Year")

```


```{r fig-yearly-maxt-compare, out.height=200, fig.cap="Comparison of yearly average maximum temperature between AGDC and other sources"}


yearly_records %>% 
    filter(Source == "agcd") %>% 
    select(Number, year, base = maxt) %>% 
    right_join(yearly_records %>% 
                   filter(Source != "agcd") %>% 
                   select(Number, Source, year, maxt)) %>% 
    ggplot() +
    geom_point(aes(base, maxt)) +
    facet_wrap(~Source) +
    theme_bw() +
    geom_abline() +
    xlab("Yearly average maximum temperature from AGCD (C)") +
    ylab("Yearly average maximum temperature from other sources (C)")

```




## Summary

Comparing with radiation, temperatures don't have big difference among data source. 

## Reference

Lilley JM, Flohr BM, Whish JPM, Farre I, Kirkegaard JA. 2019.  Defining optimal sowing and flowering periods for canola in Australia. Field Crops Research 235, 118–128.
