---
title: A faster way to trim a raster file
author: Bangyou Zheng
date: '2017-09-17'
slug: trim-raster
categories:
  - gis
tags:
  - R
  - raster
  - htp
  - gis
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```


In the plot segmentation of image-based HTP platforms (e.g. UAV, ground-vehicle), we always need to trim a plot to remove the edge effect, then extracts the pixel values we are interesting. The operation is similar with `cutline` in `gdalwarp` command.

In R, `raster` package provides a function `cellFromPolygon` to extract cull numbers from a polygon. However, `cellFromPolygon` is slightly slow, especially for a larger file. 

For example, the wheat plot from a 20M pixel Sony camera is about 3.4 MB. I want to trim this plot by 20% in each side. 

```{r read-data}
library(raster)
# Define the polygon to trim raster file 
ply <- read.table(textConnection('
434129.53992703213589 6950697.5565780829638
434127.33717670320766 6950693.9743350874633
434126.31413929269183 6950694.6008593635634
434128.51703721459489 6950698.1830912819132
434129.53992703213589 6950697.5565780829638'))
names(ply) <- c('long', 'lat')

plot <- brick('wheat-plot.tif')

# Plot the raster and polygon
plotRGB(plot)
lines(ply$long, ply$lat, col = 'red', lwd = 3)

```

The first method is to use `cellFromPolygon` to extrct the cell index. 
 
```{r cells-polygon}
library(sp)
# Create spatial polygon
sply <- SpatialPolygons(list(Polygons(list(Polygon(ply)), ID = '1')), 
                proj4string = plot@crs)

system.time(cell_idx1 <- cellFromPolygon(plot, sply))
```

An alternative method is to use `point.in.polygon` function in sp package.

```{r new-method}
system.time({
    # Generate a grid of all x and y values in the raster file
    xy <- xyFromCell(plot, seq(1, prod(dim(plot)[1:2])))
    # Check point in the polygon
    cid <- point.in.polygon(xy[,1], xy[,2], ply[,1], ply[,2])
    # Obtain the cell index
    cell_idx2 <- seq(1, prod(dim(plot)[1:2]))[cid == 1]
})

```
The first method takes about 10 S, but the second method only takes less than 1s. The two methods generate the same results for the cell index. 

```{r comparison-metho, fig.width=4, fig.height=4}

library(tidyverse)
pd <- xy %>% 
    as.data.frame() %>% 
    tbl_df() 

pd %>%
    slice(cell_idx1[[1]]) %>% 
    ggplot() +
    geom_tile(aes(x, y), fill = 'red', alpha = 0.5) +
    geom_tile(aes(x, y), fill = 'blue', alpha = 0.5, 
              data = pd[cell_idx2,]) +
    coord_equal() + 
    theme_bw() +
    xlab('') + ylab('')
```

